\RequirePackage[l2tabu, orthodox]{nag}

\documentclass{report}

\usepackage{etex}
\usepackage{graphicx}
\usepackage{float}
\usepackage{subcaption}
\usepackage{circuitikz}
\usepackage{pgfplots}

\usepackage{mathtools}

\usepackage{verbatim}

\usepackage{microtype}
\usepackage[colorlinks=false, pdfborder={0 0 0}]{hyperref}
\usepackage[all]{hypcap}
\usepackage{todonotes}

\usepackage{algorithm,algorithmicx}
\usepackage{algpseudocode}

\usepackage{cleveref}

\input{header.tex}
\input{Qcircuit.tex}

\bibliographystyle{plain}

\newfloat{program}{ht}{lop}
%----------------------------------------------------------------------------------------------------


\begin{document}
\listoftodos 

\title{Compiling Reversible Circuits}
\author{Alex Parent}
\date{\today}
\maketitle

\begin{abstract}
  A review of reversible circuits for arithmetic and circuit optimization in the context of the implementing a reversible circuit compiler.
\end{abstract}
\tableofcontents

\chapter{Introduction}


%----------------------------------------------------------------------------------------------------
\chapter{Memory Management}

In a reversible computation memory management is somewhat complicated by the reversibility constraint.

For example consider a classical computation where we have inputs $a$, and $b$ and wish to know the output $a\land b$.
Further the values of $a$, and $b$ are not needed later in the computation.
In this case we can calculate $(a,b) \mapsto a \land b$, take the output $a\land b$, and discard the values of $a$, and $b$.
Since this function is not injective (ex. $0\land 0 = 0$ and $1 \land 0 = 0$) it is not reversible.

One way to make it reversible is to use a Toffoli gate.
The Toffoli gate implements the injective function $(a,b,c) \mapsto (a,b,a\land b \oplus c)$.
If $c$ is initialized to zero we have $(a,b,0) \mapsto (a,b,a\land b)$.
This means that we cannot simply discard the values of $a$ and $b$.

\section{Bennett}
Any irreversible function can be made reversible using the ``Bennett Method''\cite{Bennett:73}.
This result was intended to show that in principle any irreversible computation can be made reversible.


\section{Janus}
\subsection{Modification Operators}
In the Janus programming language\cite{YG:2007} modification operators are available for operations that can be done in-place.
\subsection{If Conditions}

\section{Revs}\todo{A citation for this is needed.  Get this after arxiv paper is posted.}
The approach taken by Revs is complementary to to Janus.


%----------------------------------------------------------------------------------------------------

\include{arithmetic}
\include{jcc}

\bibliography{thesis}

\end{document}
